---
title: "PSID Results"
author: "Gustave Kenedi and Louis Sirugue"
date: "`r Sys.time()`"
format:
  html:
    toc: true
    df-print: paged
    embed-resources: true
editor: visual
editor_options: 
  chunk_output_type: console
execute:
  warning: false
---

## Packages

```{r}
packages <- c("here", "tidyverse", "data.table", "readxl", "janitor", "ggthemr", "caret", "patchwork", "tidymodels","ggtext", "scico", "tidylog", "skimr")
librarian::shelf(packages, quiet = TRUE, )
rm(packages)
```

## Functions

```{r}
source(here("code_lib", "useful_functions.R"))

source(here("code_lib", "inc_fct.R"))
source(here("code_lib", "gen_ranks.R"))

source(here("code_lib", "matrix_graph_fct.R"))

source(here("code_lib", "tstsls_fct.R"))
source(here("code_lib", "caption_fct.R"))
source(here("code_lib", "indiv_pred_graph.R"))
source(here("code_lib", "prediction_graph.R"))
source(here("code_lib", "prediction_rank_graph.R"))
source(here("code_lib", "prediction_quintile_graph.R"))

source(here("code_lib", "ige_comparison_graph.R"))
source(here("code_lib", "rrc_comparison_graph.R"))
source(here("code_lib", "matrix_predicted_graph.R"))
source(here("code_lib", "aum_comparison_graph.R"))
```

## Params

```{r}
# figures
fig_height = 6
fig_width = (16/9)*fig_height
ggthemr("pale")

text_size = 14
title_size = 13

theme_update(text = element_text(size = text_size),
             plot.title = element_text(size = title_size, face = "italic"),
             legend.title = element_text(face = "italic", size = 12),
             legend.text = element_text(size = 12),
             legend.background = element_rect(fill = alpha("white", .75)),
             plot.caption = element_text(size = 10, face = "italic", color = "#7d7d7d"),
             strip.text = element_text(size = 11, face = "plain"))

alpha = 0.8
my_palette = scico(5, palette = "lajolla", begin = .2, end = .95)

my.ggsave <- function(filename = default_name(plot), height = fig_height, width = fig_width) {
  ggsave(filename = filename, height = height, width = width)
}

# no scientific notation
options(scipen=999)

# path to data
data_path <- "~/Dropbox/3resources/data/psid/"
figure_path <- "/Users/gustave/Dropbox/Apps/Overleaf/Intergenerational Income Mobility/psid_validation/figures/"
```

## Load data

```{r}
#| cache: true
#| output: false

# PSID variables dictionary
psid_vars <- read_excel(paste0(data_path, "fims/psid.xlsx")) %>%
  clean_names() %>%
  mutate(var_id = 1:n()) %>% 
  pivot_longer(cols = -c(var_id, type, category, text, head_wife, var_count),
               names_prefix = "y",
               names_to = "year",
               values_to = "variable")
setDT(psid_vars)
# family interview id
psid_vars %>% filter(var_id == 8746) %>% count(text)
fam_id_vars <- psid_vars %>% filter(var_id == 8746) %>% pull(variable)

# family files 1968-2019
years <- c(1968:1997, seq(1999, 2019, 2))
for (i in years) {
  print(i)
  
  if (i %in% 1968:1993) {
    file_path <- paste0(data_path, "fam", i, "/FAM", i)
  } else {
    file_path <- paste0(data_path, "fam", i, "er", "/FAM", i, "ER")
  }
  
  assign(paste0("fam_", i), fread(paste0(file_path, ".csv")))
}

gc()
```

## Import intermediate datasets

```{r}
# correspondence table state-region
state_correspondance <- fread(here("data/state_correspondance.csv"))

# parents datasets
parents_sample_clean <- fread(here("data/parents_sample_clean_results.csv"))
parents_sample_clean <- parents_sample_clean %>% 
  mutate(across(where(is.character), ~ ifelse(. == "", NA, .)))
parents_sample_long <- fread(here("data/parents_sample_long.csv"))

# children datasets
child_sample_clean <- fread(here("data/child_sample_clean_results.csv"))
child_sample_clean <- child_sample_clean %>%
  left_join(state_correspondance, by = c("state" = "state_code"))
child_sample_long <- fread(here("data/child_sample_long.csv"))

# link parent-child
child_first_parents <- fread(here("data", "child_first_parents.csv")) %>% 
  rename(ind_id_68_child = child_fam) %>% 
  select(ind_id_68, ind_id_68_child, rel_to_head, gender)

# income data
inc_all <- fread(here("data/inc_all.csv"))

# regions
regions = child_sample_clean %>% distinct(region) %>% filter(!is.na(.)) %>% pull()
```

# Benchmark results

Note that: "In the PSID, the household head is recorded as having zero labor income if their income was actually zero or if their labor income is missing, so one cannot cleanly distinguish true zeroes with labor income. All of the main analysis only uses years of non-zero income when constructing time averages of income." ([Mazumder, 2016, p. 101](zotero://select/library/items/IGV8QXK7)) ([pdf](zotero://open-pdf/library/items/CIE4Y835?page=19&annotation=RE8H2T4M))

The `inc_fct()` function has a `zero_to_na` parameter which if set to `T` replaces 0 incomes by missing. Since the documentation isn't so clear about what zeros actually correspond, in the baseline results, zeros are not replaced, but as a robustness we drop zero income observations.

## Parameters

```{r}
# income definitions
inc_vars = c("labor income", "family labor income", "family tax income", "family total income")

# minimum and maximum child/parent age at income observation
child_age_min = 30
child_age_max = 50
parent_age_min = 30
parent_age_max = 50

# restriction for number of parent/child income observations
n_parent_inc = 1
n_child_inc = 1

# replace zero incomes
replace_zero_incomes = F
```

## Income variable creation

```{r}
child_sample_clean %>% names

# drop income variables from child and parent data.frames
child_sample_clean <- child_sample_clean %>% 
  select(-contains(c("income", "mean_age")))
parents_sample_clean <- parents_sample_clean %>% 
  select(-contains(c("income", "mean_age")))

# obtain children income variables
child_sample_clean <- child_sample_clean %>%
  left_join(inc_vars %>% 
              map(~ inc_fct(dt = child_sample_long,
                            inc_var = .x,
                            age_range = child_age_min:child_age_max,
                            zero_to_na = replace_zero_incomes) %>% 
                    select(-contains("age"))) %>% 
              reduce(full_join, by = "ind_id_68"))
child_sample_clean %>% names

# obtain parent income variables
parents_sample_clean <- parents_sample_clean %>% 
  left_join(inc_vars %>% 
              map(~ inc_fct(dt = parents_sample_long,
                            inc_var = .x,
                            age_range = parent_age_min:parent_age_max,
                            zero_to_na = replace_zero_incomes) %>% 
                    select(-contains("age"))) %>%
              reduce(full_join, by = "ind_id_68"))
parents_sample_clean %>% names

# obtain children's id
parents_sample_clean_withchildren <- parents_sample_clean %>% 
  left_join(child_first_parents %>% select(ind_id_68, ind_id_68_child))

# for each child, get one parent income measure
# each parent income measure has 2 definitions: one accounts for family size (i.e, number of head and "wife"), the other doesn't
# this is because not accounting for the number of "parents" could unreasonably affect single-headed households
child_sample_clean <- child_sample_clean %>% 
  left_join(parents_sample_clean_withchildren %>%
              mutate(across(starts_with("n_") & !contains("_n_"), ~ ifelse(gender == "Female", .x, NA), .names = "n_mother_{.col}"),
                     across(starts_with("n_") & !contains("_n_"), ~ ifelse(gender == "Male", .x, NA), .names = "n_father_{.col}")) |>
              group_by(ind_id_68_child) %>%
              summarise(presence_father = sum(gender == "Male"),
                        n_parents = n_distinct(ind_id_68),
                        across(starts_with("mean") & contains("income"), ~ mean(.x), .names = "parent_mean_{.col}"),
                        across(mean_labor_income_real_3050, ~ sum(.x), .names = "parent_sum_{.col}"),
                        across(contains(c("father", "mother")), ~ ifelse(!is.na(mean(.x, na.rm = T)), mean(.x, na.rm = T), 0))) %>% 
              rename(ind_id_68 = ind_id_68_child))

# child income variables
child_inc_vars <- child_sample_clean %>% 
  select(contains("income") & !starts_with(c("n_", "parent_"))) %>% 
  names()
child_inc_vars_labs <- paste0(c("Individual labor income\n",
                                "Family labor income",
                                "Family labor income\n(div. number of adults)",
                                "Family taxable income",
                                "Family taxable income\n(div. number of adults)",
                                "Family total income",
                                "Family total income\n(div. number of adults)"),
                              " (", child_age_min, "-", child_age_max, ")")

# parent income variables
parent_inc_vars <- child_sample_clean %>% 
  select(contains("income") & !starts_with("n_") & starts_with("parent")) %>% 
  names()
# parent_inc_vars <- parent_inc_vars[-1] # drop individual 
parent_inc_vars_labs <- paste0(c("Individual labor income (mean)",
                                 "Family labor income",
                                 "Family labor income\n(div. number of adults)",
                                 "Family taxable income",
                                 "Family taxable income\n(div. number of adults)",
                                 "Family total income",
                                 "Family total income\n(div. number of adults)",
                                 "Individual labor income (sum)"),
                               " (", parent_age_min, "-", parent_age_max, ")")

# child and parent income vectors
child_inc_vars_all <- crossing(child_inc_vars, parent_inc_vars)$child_inc_vars
parent_inc_vars_all <- crossing(child_inc_vars, parent_inc_vars)$parent_inc_vars

# define core sample as the sample of children for whom baseline household-level child and parent income variables are observed
child_sample_clean <- child_sample_clean %>%
  mutate(core_sample = ifelse(!is.na(mean_family_total_income_real_3050) & !is.na(parent_mean_mean_family_total_income_real_3050), T, F))
child_sample_clean |> count(core_sample)

# identify children whose parents have more than n_parent_inc income observations
child_sample_clean <- child_sample_clean |> 
  mutate(more_than_n_parent_inc =
           ((n_parents == 1 & presence_father == T &  n_father_n_family_total_income_real_3050 >= n_parent_inc) |
              (n_parents == 1 & presence_father == F & n_mother_n_family_total_income_real_3050 >= n_parent_inc) |
              (n_parents == 2 & n_father_n_family_total_income_real_3050 >= n_parent_inc & n_mother_n_family_total_income_real_3050 >= n_parent_inc)))
child_sample_clean |> count(more_than_n_parent_inc)

# dummy denoting children with more than n_child_inc income observations
child_sample_clean <- child_sample_clean |> 
  mutate(more_than_child_inc = (!is.na(n_family_total_income_real_3050) & n_family_total_income_real_3050 >= n_child_inc))
child_sample_clean |> count(more_than_child_inc == T)

# rank all child and parent income variables only for subset of children with all child income variables and all parent income variables
child_sample_clean <- child_sample_clean %>% 
  left_join(c(child_inc_vars, parent_inc_vars) |> 
              map(~ gen_ranks(dt = child_sample_clean |>
                                filter(core_sample == T & more_than_n_parent_inc == T & more_than_child_inc == T),
                              inc = .x)) %>%
              reduce(left_join, by ="ind_id_68"))
child_sample_clean %>% names()

# obtain log incomes and replace <=0s values by NA
child_sample_clean <- child_sample_clean |> 
  mutate(across(contains(c("mean", "sum")) & contains("income") & !contains("rank"), ~ ifelse(.x <= 0, NA, log(.x)), .names = "log_{.col}"))

# baseline core sample
dt_reg <- child_sample_clean |> filter(core_sample == T)
```

## IGE - ALL

```{r}
# estimate log-log relationship (including birth year FEs)
ige_coefs <- map2_dfr(.x = child_inc_vars_all, .y = parent_inc_vars_all,
                      ~ tidy(lm(formula(paste0("log_", .x, " ~ log_", .y, " + factor(birth_year)")), data = dt_reg)) %>% 
                        filter(grepl("log", term)) %>% 
                        mutate(child_inc_var = .x,
                               child_inc_var_lab = child_inc_vars_labs[which(child_inc_vars == .x)],
                               parent_inc_var_lab = parent_inc_vars_labs[which(parent_inc_vars == .y)],
                               num_obs = nobs(lm(formula(paste0("log_", .x, " ~ log_", .y, " + factor(birth_year)")), data = dt_reg))) %>% 
                        select(child_inc_var, child_inc_var_lab, parent_inc_var = term, parent_inc_var_lab, coef = estimate, se = std.error, num_obs))
fwrite(ige_coefs, here("data", "figures_raw_data", "ige_coefs.csv"))

# graph of baseline IGEs
ige_coefs %>%
  ggplot(aes(x = child_inc_var_lab, y = parent_inc_var_lab, fill = coef, label = paste0(round(coef, 3), "\n(", round(se, 3), ")"))) +
  geom_tile() +
  geom_text(color = "white") +
  labs(x = "Child income definition",
       y = "Parent income definition",
       caption = paste0("Naive standard errors in parenthesis. Number of children varies by income defintion since the number of negative or zero incomes varies.")
  ) +
  scale_x_discrete(guide = guide_axis(n.dodge = 2), expand = c(0,0)) +
  scale_y_discrete(expand = c(0,0)) +
  scale_fill_scico(palette = "lajolla", begin = .2, end = .95) +
  theme(legend.position = "none",
        panel.grid = element_blank())
my.ggsave(paste0(figure_path, "benchmark_ige.pdf"))
```

## RRC - ALL

```{r}
# estimate rank-rank correlations (including birth year FEs)
rrc_coefs <- map2_dfr(.x = child_inc_vars_all, .y = parent_inc_vars_all,
                      ~ tidy(lm(formula(paste0("rank_", .x, "~ rank_", .y, " + factor(birth_year)")), data = dt_reg)) %>% 
                        filter(grepl("rank", term)) %>% 
                        mutate(child_inc_var = .x,
                               child_inc_var_lab = child_inc_vars_labs[which(child_inc_vars == .x)],
                               parent_inc_var_lab = parent_inc_vars_labs[which(parent_inc_vars == .y)],
                               num_obs = nobs(lm(formula(paste0("rank_", .x, "~ rank_", .y, " + factor(birth_year)")), data = dt_reg))) %>% 
                        select(child_inc_var, child_inc_var_lab, parent_inc_var = term, parent_inc_var_lab, coef = estimate, se = std.error, num_obs))
fwrite(rrc_coefs, here("data", "figures_raw_data", "rrc_coefs.csv"))

# graph of baseline RRCs
n_obs <- rrc_coefs |> pull(num_obs) |> unique() |> formatC(format = "d", big.mark = ",")
rrc_coefs %>%
  ggplot(aes(x = child_inc_var_lab, y = parent_inc_var_lab, fill = coef, label = paste0(round(coef, 3), "\n(", round(se, 3), ")"))) +
  geom_tile() +
  geom_text(color = "white") +
  labs(x = "Child income definition",
       y = "Parent income definition",
       caption = paste0("Naive standard errors in parenthesis. Number of children: ", n_obs, ".")) +
  scale_x_discrete(guide = guide_axis(n.dodge = 2), expand = c(0,0)) +
  scale_y_discrete(expand = c(0,0)) +
  scale_fill_scico(palette = "lajolla", begin = .2, end = .95) +
  theme(legend.position = "none",
        panel.grid = element_blank())
my.ggsave(paste0(figure_path, "benchmark_rrc.pdf"))
```

## Matrix - ALL

```{r}
# estimate transition matrices
matrix <- 
  map2_dfr(.x = child_inc_vars_all, .y = parent_inc_vars_all, 
           ~ dt_reg |> 
             select(ind_id_68, all_of(paste0("rank_", c(.x, .y)))) |> 
             mutate(across(paste0("rank_", c(.x, .y)), ~ case_when(. %in% 1:20 ~ "Bottom 20%",
                                                                   . %in% 21:40 ~ "Quintile 2",
                                                                   . %in% 41:60 ~ "Quintile 3",
                                                                   . %in% 61:80 ~ "Quintile 4",
                                                                   . %in% 81:100 ~ "Top 20%"),
                           .names = "{.col}_quintile"),
                    across(paste0("rank_", c(.x, .y), "_quintile"), ~ factor(., levels = c("Bottom 20%", "Quintile 2", "Quintile 3", "Quintile 4", "Top 20%"))),
                    child_inc_var = .x,
                    child_inc_var_lab = child_inc_vars_labs[which(child_inc_vars == .x)],
                    parent_inc_var = .y,
                    parent_inc_var_lab = parent_inc_vars_labs[which(parent_inc_vars == .y)]) |> 
             select(ind_id_68, child_inc_var, child_inc_var_lab, parent_inc_var, parent_inc_var_lab,
                    child_inc_rank = paste0("rank_", .x), parent_inc_rank = paste0("rank_", .y),
                    child_inc_quintile = paste0("rank_", .x, "_quintile"), parent_inc_quintile = paste0("rank_", .y, "_quintile")))
fwrite(matrix, here("data", "figures_raw_data", "matrix_coefs.csv"))

matrix_graph_fct(matrix, facet = T)
my.ggsave(paste0(figure_path, "benchmark_matrix.pdf"), height = fig_height*3, width = fig_width*2)

matrix_gp <- matrix |> 
  group_by(child_inc_var, child_inc_var_lab, parent_inc_var, parent_inc_var_lab, parent_inc_quintile, child_inc_quintile) %>% 
  summarise(n_obs = n()) %>% 
  group_by(child_inc_var, child_inc_var_lab, parent_inc_var, parent_inc_var_lab, parent_inc_quintile) %>% 
  mutate(pct = n_obs/sum(n_obs)) %>% 
  ungroup()
fwrite(matrix_gp, here("data", "figures_raw_data", "matrix_gp_coefs.csv"))
```

## RRC - Baseline CEF

```{r}
# estimates of rank-rank correlations (including birth year FEs)
rrc_coefs <- fread(here("data", "figures_raw_data", "rrc_coefs.csv"))
rrc_coefs <- rrc_coefs |> 
  filter(child_inc_var_lab == "Family total income (30-50)" & parent_inc_var_lab == "Family total income (30-50)")

# graph of baseline RRC CEF
n_obs <- rrc_coefs |> pull(num_obs) |> unique() |> formatC(format = "d", big.mark = ",")

rrc_cef_baseline <- child_sample_clean |>
  filter(core_sample == T) %>% 
  select(rank_child = rank_mean_family_total_income_real_3050,
         rank_parent = rank_parent_mean_mean_family_total_income_real_3050) |> 
  group_by(rank_parent) %>% 
  summarise(mean_rank_child = mean(rank_child)) %>% 
  ggplot(aes(x = rank_parent, y = mean_rank_child)) +
  geom_point(alpha = .8) +
  geom_text(data = rrc_coefs, aes(x = 1, y = 90, label = paste0("RRC: ", round(coef, 3), " (", round(se, 3), ")")), hjust = 0, size = 5) +
  geom_smooth(method = "lm", se = FALSE) +
  scale_y_continuous(lim = c(0,100), expand = c(0, 0)) +
  scale_x_continuous(lim = c(0,100), expand = c(0.01, 0.01)) +
  labs(x = "Parent income rank",
       y = "Mean child income rank",
       title = "Benchmark Rank-Rank Correlation",
       # subtitle = "Parent income: family total pretax income 30-50 | Child income: family total pretax income 30-50",
       # caption = paste0("Naive standard errors in parenthesis. Number of children: ", n_obs)
  ) +
  theme(legend.position = "none")
rrc_cef_baseline
my.ggsave(paste0(figure_path, "benchmark_rrc_family.pdf"))
```

## Matrix - Baseline

```{r}
# estimate transition matrices
matrix <- fread(here("data", "figures_raw_data", "matrix_coefs.csv"))
matrix <- matrix |> 
  filter(child_inc_var_lab == "Family total income (30-50)" & parent_inc_var_lab == "Family total income (30-50)")

matrix_baseline <- matrix_graph_fct(matrix) +
  scale_x_discrete(expand = c(0,0)) +
  labs(title = "Benchmark Transition Matrix") +
  theme(legend.position = "right",
        panel.grid.major = element_blank(),
        legend.key = element_rect(linewidth = 5),
        legend.key.size = unit(1, 'lines'),
        legend.key.height = unit(2, "cm"),
        legend.key.width = unit(1, "cm"))
matrix_baseline
my.ggsave(paste0(figure_path, "benchmark_matrix_family.pdf"), height = fig_height*2, width = fig_width*1.75)
```

## RRC + Matrix - Baseline

```{r}
patchwork <- rrc_cef_baseline + matrix_baseline + plot_annotation(caption = paste0("Naive standard errors in parenthesis. Number of children: ", n_obs, "\nParent income: family total pretax income (avg. 30-50 yrs old) | Child income: family total pretax income (avg. 30-50 yrs old)"))
patchwork +
  plot_annotation(tag_levels = "A")
my.ggsave(paste0(figure_path, "benchmark_estimates.pdf"))
```

## RRC and AUM by Region

```{r}
# baseline child and parent income definitions
child_inc_reg = "mean_family_total_income_real_3050"
parent_inc_reg = "parent_mean_mean_family_total_income_real_3050"

# estimate rank-rank correlations (including birth year FEs)
rrc_coefs <- pmap_dfr(list(x = rep(child_inc_reg, length(regions)), 
                           y = rep(parent_inc_reg, length(regions)),
                           z = regions),
                      ~ tidy(lm(formula(paste0("rank_", ..1, "~ rank_", ..2, " + factor(birth_year)")), data = dt_reg |> filter(region == ..3))) %>%
                        filter(grepl("rank|(Intercept)", term)) |> 
                        mutate(region = ..3,
                               child_inc_var = ..1,
                               child_inc_var_lab = child_inc_vars_labs[which(child_inc_vars == ..1)],
                               parent_inc_var_lab = parent_inc_vars_labs[which(parent_inc_vars == ..2)],
                               num_obs = nobs(lm(formula(paste0("rank_", ..1, "~ rank_", ..2, " + factor(birth_year)")), data = dt_reg |> filter(region == ..3)))) %>% 
                        select(region, child_inc_var, child_inc_var_lab, parent_inc_var = term, parent_inc_var_lab, coef = estimate, se = std.error, num_obs))
fwrite(rrc_coefs, here("data/figures_raw_data/rrc_coefs_region.csv"))

# compute AUM
rrc_coefs <- rrc_coefs |> 
  group_by(region) |> 
  mutate(aum = coef[parent_inc_var == "(Intercept)"] + 25 * coef[parent_inc_var == paste0("rank_", parent_inc_reg)]) |> 
  filter(parent_inc_var != "(Intercept)")

# number of observations clean
rrc_coefs <- rrc_coefs |>
  mutate(num_obs =  formatC(num_obs, format = "d", big.mark = ","))

# graph CEF
rrc_cef_baseline <- dt_reg %>% 
  filter(!is.na(region)) |>
  select(region,
         rank_child = paste0("rank_", child_inc_reg),
         rank_parent = paste0("rank_", parent_inc_reg)) |> 
  group_by(region, rank_parent) %>% 
  summarise(mean_rank_child = mean(rank_child)) %>% 
  ggplot(aes(x = rank_parent, y = mean_rank_child)) +
  geom_point(alpha = .8) +
  geom_text(data = rrc_coefs, aes(x = 1, y = 93, label = paste0("RRC: ", round(coef, 3), " (", round(se, 3), ")")), hjust = 0, size = 5) +
  geom_text(data = rrc_coefs, aes(x = 1, y = 81, label = paste0("AUM: ", round(aum, 2))), hjust = 0, size = 5) +
  geom_richtext(data = rrc_coefs, aes(x = 100, y = 5, label = paste0("<span style='color:#78695F; font-size:10pt'>*Number of observations: ", num_obs, "*</span>")), hjust = 1, color = "transparent", fill = alpha("white", .5), size = 5) +
  geom_smooth(method = "lm", se = FALSE) +
  facet_wrap(vars(factor(region, levels = c("South", "Midwest", "West", "Northeast"))), ncol = 2) +
  scale_y_continuous(lim = c(0,100), expand = c(0, 0)) +
  scale_x_continuous(lim = c(0,100), expand = c(0.01, 0.01)) +
  labs(x = "Parent income rank",
       y = "Mean child income rank",
       # title = "Benchmark Rank-Rank Correlation and Absolute Upward Mobility",
       caption = "Parent income: family total pretax income (avg. 30-50 yrs old) | Child income: family total pretax income (avg. 30-50 yrs old") +
  theme(legend.position = "none",
        strip.text = element_text(size = 14, face = "bold"),
        panel.spacing.x = unit(2, "lines"))
rrc_cef_baseline
my.ggsave(paste0(figure_path, "benchmark_geo_family.pdf"))
```

### Robustness

#### Only SRC Sample

##### Rank variable creation

```{r}
# obtain benchmark RRCs
child_sample_clean <- child_sample_clean %>% 
  select(-contains("rank"))

child_sample_clean <- child_sample_clean %>% 
  left_join(c(child_inc_vars, parent_inc_vars) |> 
              map(~ gen_ranks(dt = child_sample_clean |> 
                                filter(core_sample == T & itw_num_cat == "Member of, or moved into, a family from the 1968 SRC cross-section sample"),
                              inc = .x)) %>%
              reduce(left_join, by ="ind_id_68"))
child_sample_clean %>% names()

dt_reg <- child_sample_clean |> filter(core_sample == T & itw_num_cat == "Member of, or moved into, a family from the 1968 SRC cross-section sample")
```

##### RRC + Matrix - Baseline

```{r}
# estimate rank-rank correlations (including birth year FEs)
rrc_coefs <- map2_dfr(.x = "mean_family_total_income_real_3050", .y = "parent_mean_mean_family_total_income_real_3050",
                      ~ tidy(lm(formula(paste0("rank_", .x, "~ rank_", .y, " + factor(birth_year)")), data = dt_reg)) %>% 
                        filter(grepl("rank", term)) %>% 
                        mutate(child_inc_var = .x,
                               child_inc_var_lab = child_inc_vars_labs[which(child_inc_vars == .x)],
                               parent_inc_var_lab = parent_inc_vars_labs[which(parent_inc_vars == .y)],
                               num_obs = nobs(lm(formula(paste0("rank_", .x, "~ rank_", .y, " + factor(birth_year)")), data = dt_reg))) %>% 
                        select(child_inc_var, child_inc_var_lab, parent_inc_var = term, parent_inc_var_lab, coef = estimate, se = std.error, num_obs))

# graph of baseline RRC CEF
n_obs <- rrc_coefs |> pull(num_obs) |> unique() |> formatC(format = "d", big.mark = ",")

rrc_cef_baseline <- dt_reg %>% 
  select(rank_child = rank_mean_family_total_income_real_3050,
         rank_parent = rank_parent_mean_mean_family_total_income_real_3050) |> 
  group_by(rank_parent) %>% 
  summarise(mean_rank_child = mean(rank_child)) %>% 
  ggplot(aes(x = rank_parent, y = mean_rank_child)) +
  geom_point(alpha = .8) +
  geom_text(data = rrc_coefs, aes(x = 1, y = 90, label = paste0("RRC: ", round(coef, 3), " (", round(se, 3), ")")), hjust = 0, size = 5) +
  geom_smooth(method = "lm", se = FALSE) +
  scale_y_continuous(lim = c(0,100), expand = c(0, 0)) +
  scale_x_continuous(lim = c(0,100), expand = c(0.01, 0.01)) +
  labs(x = "Parent income rank",
       y = "Mean child income rank",
       title = "Benchmark Rank-Rank Correlation",
       # subtitle = "Parent income: family total pretax income 30-50 | Child income: family total pretax income 30-50",
       # caption = paste0("Naive standard errors in parenthesis. Number of children: ", n_obs)
  ) +
  theme(legend.position = "none")
rrc_cef_baseline

# estimate transition matrices
matrix <- 
  map2_dfr(.x = "mean_family_total_income_real_3050", .y = "parent_mean_mean_family_total_income_real_3050", 
           ~ dt_reg |> 
             select(ind_id_68, all_of(paste0("rank_", c(.x, .y)))) |> 
             mutate(across(paste0("rank_", c(.x, .y)), ~ case_when(. %in% 1:20 ~ "Bottom 20%",
                                                                   . %in% 21:40 ~ "Quintile 2",
                                                                   . %in% 41:60 ~ "Quintile 3",
                                                                   . %in% 61:80 ~ "Quintile 4",
                                                                   . %in% 81:100 ~ "Top 20%"),
                           .names = "{.col}_quintile"),
                    across(paste0("rank_", c(.x, .y), "_quintile"), ~ factor(., levels = c("Bottom 20%", "Quintile 2", "Quintile 3", "Quintile 4", "Top 20%"))),
                    child_inc_var = .x,
                    child_inc_var_lab = child_inc_vars_labs[which(child_inc_vars == .x)],
                    parent_inc_var = .y,
                    parent_inc_var_lab = parent_inc_vars_labs[which(parent_inc_vars == .y)]) |> 
             select(ind_id_68, child_inc_var, child_inc_var_lab, parent_inc_var, parent_inc_var_lab,
                    child_inc_rank = paste0("rank_", .x), parent_inc_rank = paste0("rank_", .y),
                    child_inc_quintile = paste0("rank_", .x, "_quintile"), parent_inc_quintile = paste0("rank_", .y, "_quintile")))

matrix_baseline <- matrix |> 
  group_by(child_inc_var, child_inc_var_lab, parent_inc_var, parent_inc_var_lab, parent_inc_quintile, child_inc_quintile) %>% 
  summarise(n_obs = n()) %>% 
  group_by(child_inc_var, child_inc_var_lab, parent_inc_var, parent_inc_var_lab, parent_inc_quintile) %>% 
  mutate(pct = n_obs/sum(n_obs)) %>% 
  ungroup() |> 
  ggplot(aes(x = parent_inc_quintile, y = pct, fill = fct_rev(child_inc_quintile))) +
  geom_col(position = "fill") +
  geom_text(aes(y = pct, label = paste0(round(pct*100,2), "%")), position = position_fill(vjust = .5),
            show.legend = FALSE, color = "white", size = 4) +
  scale_fill_manual(values = alpha(my_palette[c(5,4,3,2,1)], alpha)) +
  scale_y_continuous(expand = c(0,0), labels = scales::percent) +
  scale_x_discrete(expand = c(0,0)) +
  labs(x = "Parent income quintile",
       y = NULL,
       title = "Benchmark Transition Matrix",
       fill = "Child income\nquintile:") +
  theme(legend.position = "right",
        panel.grid.major = element_blank(),
        legend.key = element_rect(linewidth = 5),
        legend.key.size = unit(1, 'lines'),
        legend.key.height = unit(2, "cm"),
        legend.key.width = unit(1, "cm"))
matrix_baseline

patchwork <- rrc_cef_baseline + matrix_baseline + plot_annotation(caption = paste0("Naive standard errors in parenthesis. Number of children: ", n_obs, "\nParent income: family total pretax income (avg. 30-50 yrs old) | Child income: family total pretax income (avg. 30-50 yrs old)"))
patchwork +
  plot_annotation(tag_levels = "A")
my.ggsave(paste0(figure_path, "benchmark_estimates_src_sample.pdf"))
```

# TSTSLS Results

## Setup

```{r}
child_sample_clean <- child_sample_clean %>% 
  select(-contains(c("income", "mean_age")))
parents_sample_clean <- parents_sample_clean %>% 
  select(-contains(c("income", "mean_age")))

# convert instruments to factors
parents_sample_clean <- parents_sample_clean %>% 
  mutate(state = as.character(state),
         birth_year = as.character(birth_year),
         race = as.character(race))

# full set of instruments to keep sample constant
dem_chars <- c("birth_year", "race")
full_instru = c("edu_level", "occ_3_digit_3050", dem_chars, "state")

# instruments
instru = list(full_instru[1], full_instru[1:2], full_instru[1:4], full_instru[1:5])
# instruments labels
full_instru_labs <- c("education", "occupation (3-digit)", "race", "\nbirth cohort", "state FE")
instru_labs = list(full_instru_labs[1],
                   paste0(full_instru_labs[1:2], collapse = " + "),
                   paste0(full_instru_labs[1:4], collapse = " + "),
                   paste0(full_instru_labs[1:5], collapse = " + "))

# instruments labels
instru_labs_regions = rep(list(full_instru_labs[1],
                               paste0(full_instru_labs[1:2], collapse = " + "),
                               paste0(full_instru_labs[1:4], collapse = " + "),
                               paste0(full_instru_labs[1:5], collapse = " + ")), length(regions))
region_labs = rep(regions, each = length(instru))
```

## First-stage parameters

```{r}
# parameters
parent_inc = "labor income"
parent_age_min = 30
parent_age_max = 50
parent_inc_var = paste0("mean_", str_replace_all(parent_inc, " ", "_"), "_real_", parent_age_min, parent_age_max)
parent_inc_pred_def = "sum"
n_parent_inc_obs = 1

child_inc = "family total income"
child_age_min = 30
child_age_max = 50
child_inc_var = paste0("mean_", str_replace_all(child_inc, " ", "_"), "_real_", child_age_min, child_age_max)
```

### First-stage estimation

*Predictors:* - education level (7 categories) - occupation (446 cat.) - demographic characteristics (birth cohort, race (5 cat.) - location of residence characteristics -\> state FE

*Questions:* 1) should we bootstrap parents' sample as in Jerrim et al.? -\> NO 2) how should we do out of sample prediction? -\> 5-fold cross-validation 3) Should I have multiple observations for parents' with several children? -\> NO

*Notes:* 1. Lee and Solon (2009) and Mazumder (2016) restrict their analysis to the SRC (Survey Research Center) component of the PSID, i.e., the nationally-representative part. For us, it doesn't matter so much (though it quite significantly reduces the sample size) but we do this as a robustness check.

```{r}
# estimate first-stage ----
# parent sample - obtain incomes
parents_sample_clean <- parents_sample_clean %>%
  left_join(inc_fct(dt = parents_sample_long,
                    inc_var = parent_inc,
                    age_range = parent_age_min:parent_age_max,
                    zero_to_na = replace_zero_incomes) %>% 
              mutate(across(starts_with("n_"), ~ ifelse(is.na(.), 0, .))))
# child sample - obtain incomes
child_sample_clean <- child_sample_clean %>%
  filter(core_sample == T) |> 
  left_join(inc_fct(dt = child_sample_long,
                    inc_var = child_inc,
                    age_range = child_age_min:child_age_max,
                    zero_to_na = replace_zero_incomes))

# first-stage
set.seed(1234)
fs_predict <- 1:length(instru) %>% 
  map(~ tstsls_fct(dt = parents_sample_clean,
                   dt_child = child_sample_clean,
                   instruments = instru[[.x]],
                   constant_instruments = NULL,
                   parent_inc_var = parent_inc_var,
                   replace_0_income = NULL,
                   fs_log_inc = T,
                   parent_inc_pred_def = parent_inc_pred_def,
                   child_inc_var = child_inc_var,
                   n_parent_inc_obs = n_parent_inc_obs))
# ----
```

#### Out-of-sample predictions

##### Individual-level

```{r}
# prediction vs observed individual income graph ----
indiv_pred_graphs <- 1:length(instru) %>% 
  map(~ indiv_pred_graph(dt = fs_predict[[.x]][[9]],
                         caption_lab = caption_fct(fs = fs_predict[[.x]][[8]],
                                                   instru_labels = instru_labs[[.x]])))
wrap_plots(map(indiv_pred_graphs, 1))
ggsave(paste0(figure_path, "baseline_indiv_prediction_graphs.pdf"),
       height = fig_height, width = fig_width)

wrap_plots(map(indiv_pred_graphs, 2))
ggsave(paste0(figure_path, "baseline_indiv_prediction_rank_graphs.pdf"),
       height = fig_height, width = fig_width)

wrap_plots(map(indiv_pred_graphs, 3), guides = "collect", nrow = 2) & theme(legend.position = "top") 
ggsave(paste0(figure_path, "baseline_indiv_prediction_quintile_graphs.pdf"),
       height = fig_height*1.5, width = fig_width)

# export data for paper graphs
indiv_pred_graphs <- 1:length(instru) %>% 
  map(~ indiv_pred_graph(dt = fs_predict[[.x]][[9]],
                         by_gender = F,
                         caption_lab = caption_fct(fs = fs_predict[[.x]][[8]],
                                                   instru_labels = instru_labs[[.x]])))
fwrite(indiv_pred_graphs[[4]][[4]], here("data/figures_raw_data/baseline_indiv_prediction_rank_graphs.csv"))
# ----
```

##### Household-level

```{r}
# prediction vs observed family income graph ----
graphs_prediction <- 1:length(instru) %>% 
  map(~ prediction_graph(dt = fs_predict[[.x]][[1]],
                         caption_lab = caption_fct(fs = fs_predict[[.x]][[8]],
                                                   instru_labels = instru_labs[[.x]])))
wrap_plots(graphs_prediction, nrow = 2)
ggsave(paste0(figure_path, "baseline_family_prediction_graphs.pdf"),
       height = fig_height, width = fig_width)
# ----

# prediction vs observed income rank graph ----
graphs_prediction_rank <- 1:length(instru) %>% 
  map(~ prediction_rank_graph(dt = fs_predict[[.x]][[1]],
                              caption_lab = caption_fct(fs = fs_predict[[.x]][[8]],
                                                        instru_labels = instru_labs[[.x]])))
wrap_plots(graphs_prediction_rank, nrow = 2)
ggsave(paste0(figure_path, "baseline_family_prediction_rank_graphs.pdf"),
       height = fig_height, width = fig_width)
# ----

# prediction vs observed income quintile graph ----
graphs_quintile <- 1:length(instru) %>% 
  map(~ prediction_quintile_graph(dt = fs_predict[[.x]][[1]],
                                  caption_lab = caption_fct(fs = fs_predict[[.x]][[8]],
                                                            instru_labels = instru_labs[[.x]])))
wrap_plots(graphs_quintile, guides = "collect", nrow = 2) & theme(legend.position = "top")
ggsave(paste0(figure_path, "baseline_family_prediction_quintile_graphs.pdf"),
       height = fig_height*1.5, width = fig_width)
# ----
```

## National-level results

### IGE - All FS

```{r}
# IGE comparison graph ----
graphs_ige <- 1:length(instru) %>% 
  map(~ ige_comparison_graph(dt = fs_predict[[.x]][[1]],
                             caption_lab = caption_fct(fs = fs_predict[[.x]][[8]],
                                                       instru_labels = instru_labs[[.x]]),
                             title_size = title_size,
                             text_size = text_size,
                             annotation_size = 5))
wrap_plots(graphs_ige, nrow = 2)
ggsave(paste0(figure_path, "baseline_ige_comp_graphs.pdf"),
       height = fig_height, width = fig_width)
# ----
```

### RRC - All FS

```{r}
# RRC comparison graph ----
graphs_rrc <- 1:length(instru) %>% 
  map(~ rrc_comparison_graph(dt = fs_predict[[.x]][[1]],
                             caption_lab = caption_fct(fs = fs_predict[[.x]][[8]],
                                                       instru_labels = instru_labs[[.x]]),
                             
                             title_size = title_size,
                             text_size = text_size,
                             annotation_size = 5))
wrap_plots(graphs_rrc, nrow = 2)
ggsave(paste0(figure_path, "baseline_rrc_comp_graphs.pdf"),
       height = fig_height, width = fig_width)

# export underlying data
data_rrc <- fs_predict[[4]][[1]]
write_rds(data_rrc, here("data/figures_raw_data/data_rrc.rds"))
# ----
```

### Matrix - All FS

```{r}
# transition matrix comparison graph ----
full_instru_labs_new <- c("education", "occupation (3-digit)", "race", "\nbirth cohort", "state FE")
instru_labs_new = list(full_instru_labs_new[1],
                       paste0(full_instru_labs_new[1:2], collapse = " + "),
                       paste0(full_instru_labs_new[1:4], collapse = " + "),
                       paste0(full_instru_labs_new[1:5], collapse = " + "))
graphs_matrix <- 1:length(instru) %>% 
  map(~ matrix_predicted_graph(dt = fs_predict[[.x]][[1]],
                               caption_lab = caption_fct(fs = fs_predict[[.x]][[8]],
                                                         instru_labels = instru_labs_new[[.x]]),
                               title_size = 8,
                               ols_legend_none = T,
                               tstsls_legend_none = T,
                               pct_size = 2.5,
                               n_dodge = 2))
wrap_plots(map(graphs_matrix, 1), nrow = 2)
# graphs_matrix[[1]][[2]] + (wrap_plots(map(graphs_matrix, 3), nrow = 2, guides = "collect") & theme(legend.position = 'top')) + plot_layout(widths = c(.3, .7))
ggsave(paste0(figure_path, "baseline_matrix_comp_graphs.pdf"),
       height = fig_height*1.5, width = fig_width)
# ----
```

### RRC - Baseline

```{r}
baseline_rrc <- graphs_rrc[[4]]
baseline_rrc
my.ggsave(paste0(figure_path, "baseline_rrc.pdf"))
```

### Transition matrices - Baseline

```{r}
baseline_matrix <- matrix_predicted_graph(dt = fs_predict[[4]][[1]],
                                          caption_lab = caption_fct(fs = fs_predict[[4]][[8]],
                                                                    instru_labels = instru_labs_new[[4]]),
                                          title_size = title_size,
                                          ols_legend_none = T,
                                          tstsls_legend_none = F,
                                          legend_size = 12,
                                          pct_size = 4,
                                          n_dodge = 1)

baseline_matrix_tstsls <- baseline_matrix[[3]] +
  labs(title = "TSTSLS") +
  guides(fill = guide_legend(reverse = F)) +
  theme(legend.key = element_rect(linewidth = 5),
        legend.key.size = unit(1, 'lines'),
        legend.key.height = unit(2, "cm"),
        legend.key.width = unit(1, "cm"))
baseline_matrix_ols <- baseline_matrix[[2]]
baseline_matrix_graph <- baseline_matrix_ols + baseline_matrix_tstsls
baseline_matrix_graph
my.ggsave(paste0(figure_path, "baseline_matrix.pdf"))
```

### RRC + Transition Matrix - Baseline

```{r}
baseline_matrix <- matrix_predicted_graph(dt = fs_predict[[4]][[1]],
                                          caption_lab = caption_fct(fs = fs_predict[[4]][[8]],
                                                                    instru_labels = instru_labs_new[[4]]),
                                          title_size = title_size,
                                          ols_legend_none = F,
                                          tstsls_legend_none = F,
                                          legend_size = 10,
                                          pct_size = 3,
                                          n_dodge = 2)

patchwork <- (baseline_rrc + labs(title = "Rank-Rank Correlation")) +
  (((baseline_matrix[[2]] + guides(fill = guide_legend(reverse = F))) /
      (baseline_matrix[[3]] + labs(title = "TSTSLS")  + guides(fill = guide_legend(reverse = F))) +
      plot_layout(guides = "collect")) &
     theme(legend.position = "right",
           legend.key = element_rect(linewidth = 5),
           legend.key.size = unit(1, 'lines'),
           legend.key.height = unit(2, "cm"),
           legend.key.width = unit(0.5, "cm")))
patchwork +
  plot_annotation(tag_levels = "A") + plot_layout(widths = c(.6, .5))
my.ggsave(paste0(figure_path, "benchmark_estimates_ols_tstsls.pdf"))
```

## Geographic-level results

Considering the sample size (5,088 children), we estimate intergenerational mobility only at the macro region level (4 categories).

Census Bureau statistical regions: 1. Northeast 2. Midwest 3. South 4. West.

### Setup

```{r}
instru_reg = rep(1:length(instru), length(regions))
regions_reg <- rep(regions, each = length(instru))
```

### RRC and AUM - Baseline

```{r}
# RRC comparison graph ----
rrc_x = 32
aum_x = 52

graphs_rrc <- map2(.x = instru_reg,
                   .y = regions_reg,
                   ~ rrc_comparison_graph(dt = fs_predict[[.x]][[1]] |> filter(region == .y),
                                          caption_lab = caption_fct(fs = fs_predict[[.x]][[8]],
                                                                    instru_labels = instru_labs[[.x]]),
                                          with_aum = T,
                                          rank_parent_text = c(rep(c(rrc_x, aum_x, 3), 2), rrc_x, aum_x, 100),
                                          mean_rank_child_text = c(rep(95, 2), rep(87, 3), rep(78, 3), 2),
                                          with_caption = F,
                                          title_size = 12,
                                          text_size = 10,
                                          num_obs_size = 8,
                                          annotation_size = 3))

wrap_plots(map(1:length(regions), ~ wrap_elements(wrap_plots(graphs_rrc[[4*.x]] + labs(title = NULL)) +
                                                    plot_annotation(title = paste0("Region: ", regions[.x]),
                                                                    theme = theme(plot.title = element_text(size = 11))))),
           nrow = 2)
my.ggsave(paste0(figure_path, "regions_baseline_comp_graph.pdf"))
```

### RRC and AUM - ALL

```{r}
full_instru_labs_new <- c("education", "\noccupation (3-digit)", "race", "\nbirth cohort", "state FE")
instru_labs_new = list(full_instru_labs_new[1],
                       paste0(full_instru_labs_new[1:2], collapse = " + "),
                       paste0(full_instru_labs_new[1:4], collapse = " + "),
                       paste0(full_instru_labs_new[1:5], collapse = " + "))

rrc_x = 30
aum_x = 75

graphs_rrc <- map2(.x = instru_reg,
                   .y = regions_reg,
                   ~ rrc_comparison_graph(dt = fs_predict[[.x]][[1]] |> filter(region == .y),
                                          caption_lab = caption_fct(fs = fs_predict[[.x]][[8]],
                                                                    instru_labels = instru_labs_new[[.x]]),
                                          with_aum = T,
                                          region_lab = T,
                                          rank_parent_text = c(rep(c(rrc_x, aum_x, 1), 2), rrc_x, aum_x, 100),
                                          mean_rank_child_text = c(rep(95, 2), rep(83, 3), rep(71, 3), 2),
                                          with_caption = F,
                                          title_size = 10,
                                          text_size = 10,
                                          num_obs_size = 8,
                                          annotation_size = 3))

wrap_plots(map(1:length(regions), ~ wrap_elements(wrap_plots(graphs_rrc[(4*.x-3):(4*.x)], ncol = length(instru)) +
                                                    plot_annotation(title = paste0("Region: ", regions[.x]),
                                                                    theme = theme(plot.title = element_text(size = 11))))),
           nrow = length(regions))

ggsave(paste0(figure_path, "regions_baseline_all_comp_graphs.pdf"),
       height = 2*fig_height, width = fig_width)
```

## Robustness

### Dealing with zero labor incomes in PSID

According to Mazumder (though it seems not according to the codebooks): "In the PSID, the household head is recorded as having zero labor income if their income was actually zero or if their labor income is missing, so one cannot cleanly distinguish true zeroes with labor income. All of the main analysis only uses years of non-zero income when constructing time averages of income." ([Mazumder, 2016, p. 101](zotero://select/library/items/IGV8QXK7)) ([pdf](zotero://open-pdf/library/items/CIE4Y835?page=19&annotation=RE8H2T4M))

Thus I check what happens when dropping these zero labor incomes.

#### First-stage estimation

```{r}
# estimate first-stage ----
child_sample_clean <- child_sample_clean %>% 
  select(-contains(c("income", "mean_age")))
parents_sample_clean <- parents_sample_clean %>% 
  select(-contains(c("income", "mean_age")))

replace_zero_incomes = T

# parent sample - obtain incomes
parents_sample_clean <- parents_sample_clean %>%
  left_join(inc_fct(dt = parents_sample_long,
                    inc_var = parent_inc,
                    age_range = parent_age_min:parent_age_max,
                    zero_to_na = replace_zero_incomes) %>% 
              mutate(across(starts_with("n_"), ~ ifelse(is.na(.), 0, .))))
# child sample - obtain incomes
child_sample_clean <- child_sample_clean %>%
  filter(core_sample == T) |> 
  left_join(inc_fct(dt = child_sample_long,
                    inc_var = child_inc,
                    age_range = child_age_min:child_age_max,
                    zero_to_na = replace_zero_incomes))

# first-stage
set.seed(1234)
fs_predict <- 1:length(instru) %>% 
  map(~ tstsls_fct(dt = parents_sample_clean,
                   dt_child = child_sample_clean,
                   instruments = instru[[.x]],
                   constant_instruments = NULL,
                   parent_inc_var = parent_inc_var,
                   replace_0_income = NULL,
                   fs_log_inc = T,
                   parent_inc_pred_def = parent_inc_pred_def,
                   child_inc_var = child_inc_var,
                   n_parent_inc_obs = n_parent_inc_obs))
# ----
```

#### RRC - All FS

```{r}
# RRC comparison graph ----
graphs_rrc <- 1:length(instru) %>% 
  map(~ rrc_comparison_graph(dt = fs_predict[[.x]][[1]],
                             caption_lab = caption_fct(fs = fs_predict[[.x]][[8]],
                                                       instru_labels = instru_labs[[.x]]),
                             
                             title_size = title_size,
                             text_size = text_size,
                             annotation_size = 5))
wrap_plots(graphs_rrc, nrow = 2)
ggsave(paste0(figure_path, "baseline_rrc_comp_graphs_replace_zero_to_na.pdf"),
       height = fig_height, width = fig_width)
# ----
```

#### Matrix - All FS

```{r}
# transition matrix comparison graph ----
full_instru_labs_new <- c("education", "occupation (3-digit)", "race", "\nbirth cohort", "state FE")
instru_labs_new = list(full_instru_labs_new[1],
                       paste0(full_instru_labs_new[1:2], collapse = " + "),
                       paste0(full_instru_labs_new[1:4], collapse = " + "),
                       paste0(full_instru_labs_new[1:5], collapse = " + "))
graphs_matrix <- 1:length(instru) %>% 
  map(~ matrix_predicted_graph(dt = fs_predict[[.x]][[1]],
                               caption_lab = caption_fct(fs = fs_predict[[.x]][[8]],
                                                         instru_labels = instru_labs_new[[.x]]),
                               title_size = 8,
                               ols_legend_none = T,
                               tstsls_legend_none = T,
                               pct_size = 2.5,
                               n_dodge = 2))
wrap_plots(map(graphs_matrix, 1), nrow = 2)
# graphs_matrix[[1]][[2]] + (wrap_plots(map(graphs_matrix, 3), nrow = 2, guides = "collect") & theme(legend.position = 'top')) + plot_layout(widths = c(.3, .7))
ggsave(paste0(figure_path, "baseline_matrix_comp_graphs_replace_zero_to_na.pdf"),
       height = fig_height*1.5, width = fig_width)
# ----
```

### Dividing income by number of reference person/partner

#### First-stage estimation

```{r}
# estimate first-stage ----
child_sample_clean <- child_sample_clean %>% 
  select(-contains(c("income", "mean_age")))
parents_sample_clean <- parents_sample_clean %>% 
  select(-contains(c("income", "mean_age")))

replace_zero_incomes = F
parent_inc_pred_def = "mean"
child_inc_var = paste0("mean_", str_replace_all(child_inc, " ", "_"), "_real_n_", child_age_min, child_age_max)

# parent sample - obtain incomes
parents_sample_clean <- parents_sample_clean %>%
  left_join(inc_fct(dt = parents_sample_long,
                    inc_var = parent_inc,
                    age_range = parent_age_min:parent_age_max) %>% 
              mutate(across(starts_with("n_"), ~ ifelse(is.na(.), 0, .))))
# child sample - obtain incomes
child_sample_clean <- child_sample_clean %>%
  filter(core_sample == T) |> 
  left_join(inc_fct(dt = child_sample_long,
                    inc_var = child_inc,
                    age_range = child_age_min:child_age_max))

# first-stage
set.seed(1234)
fs_predict <- 1:length(instru) %>% 
  map(~ tstsls_fct(dt = parents_sample_clean,
                   dt_child = child_sample_clean,
                   instruments = instru[[.x]],
                   constant_instruments = NULL,
                   parent_inc_var = parent_inc_var,
                   replace_0_income = NULL,
                   fs_log_inc = T,
                   parent_inc_pred_def = parent_inc_pred_def,
                   child_inc_var = child_inc_var,
                   n_parent_inc_obs = n_parent_inc_obs))
# ----
```

#### RRC - All FS

```{r}
# RRC comparison graph ----
graphs_rrc <- 1:length(instru) %>% 
  map(~ rrc_comparison_graph(dt = fs_predict[[.x]][[1]],
                             caption_lab = caption_fct(fs = fs_predict[[.x]][[8]],
                                                       instru_labels = instru_labs[[.x]]),
                             
                             title_size = title_size,
                             text_size = text_size,
                             annotation_size = 5))
wrap_plots(graphs_rrc, nrow = 2)
ggsave(paste0(figure_path, "baseline_rrc_comp_graphs_div_num_adults.pdf"),
       height = fig_height, width = fig_width)
# ----
```

#### Matrix - All FS

```{r}
# transition matrix comparison graph ----
full_instru_labs_new <- c("education", "occupation (3-digit)", "race", "\nbirth cohort", "state FE")
instru_labs_new = list(full_instru_labs_new[1],
                       paste0(full_instru_labs_new[1:2], collapse = " + "),
                       paste0(full_instru_labs_new[1:4], collapse = " + "),
                       paste0(full_instru_labs_new[1:5], collapse = " + "))
graphs_matrix <- 1:length(instru) %>% 
  map(~ matrix_predicted_graph(dt = fs_predict[[.x]][[1]],
                               caption_lab = caption_fct(fs = fs_predict[[.x]][[8]],
                                                         instru_labels = instru_labs_new[[.x]]),
                               title_size = 8,
                               ols_legend_none = T,
                               tstsls_legend_none = T,
                               pct_size = 2.5,
                               n_dodge = 2))
wrap_plots(map(graphs_matrix, 1), nrow = 2)
# graphs_matrix[[1]][[2]] + (wrap_plots(map(graphs_matrix, 3), nrow = 2, guides = "collect") & theme(legend.position = 'top')) + plot_layout(widths = c(.3, .7))
ggsave(paste0(figure_path, "baseline_matrix_comp_graphs_div_num_adults.pdf"),
       height = fig_height*1.5, width = fig_width)
# ----
```

## Session info

```{r}
sessionInfo()
```
